---
title: "Exercises 3. Simple R Functions"
author: "Lauren VanValkenburg"
date: "1/31/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1.

(a) Write functions \texttt{tmpFn1} and \texttt{tmpFn2} such that if \texttt{xVec} is the vector $(x_1, x_2, ... ,x_n)$, then \texttt{tmpFn1(xVec)} returns vector $(x_1, x_2^2, ..., x_n^n)$ and \texttt{tmpFn2(xVec)} returns the vector $(x_1, \frac{x_2^2}{2}, ..., \frac{x_n^n}{n})$.

Here is \texttt{tmpFn1}
```{r}

tmpFn1 <- function(xVec){
  return(xVec^(1:length(xVec)))
}
```

and now \texttt{tmpFn2}

```{r}
tmpFn2 <- function(xVec2)
{
  n = length(xVec2)
  return(xVec2^(1:n)/(1:n))
}
```

(b) Now write a fuction \texttt{tmpFn3} which takes 2 arguments $x$ and $n$ where $x$ is a single number and $n$ is a strictly positive integer.  The function should return the value of   
$$1 + \frac{x}{1} + \frac{x^2}{2} + \frac{x^3}{3} + ... + \frac{x^n}{n}$$
```{r}
tmpFn3 <- function(x,n)
{
  1+sum((x^(1:n))/(1:n))
}
```
******

### 2. Write a funciton \texttt{tmpFn(xVec)} such that if \texttt{xVec} is the vector $x = (x_1, ...,x_n)$ then \texttt{tmpFn(xVec)} returns the vector of moving averages:

$$\frac{x_1 + x_2 + x_3}{3}, \frac{x_2 + x_3 + x_4}{3}, ...,\frac{x_{n-2} + x_{n-1} + x_n}{3}$$
```{r}
tmpFn <- function(xVec)
{
  n <- length(xVec)
  (xVec[-c(n-1,n)]+xVec[-c(1,n)]+xVec[-c(1,2)])/3
}
```
******

### 3. Consider the continuous function

$$f(x) = \left\{\begin{matrix}
x^2 + 2x + 3 & if  & x < 0 \\ 
x + 3 & if & 0 \leq x < 2 \\ 
x^2 + 4x - 7 & if & 2 \leq  x 
\end{matrix}\right.$$

Write a function \texttt{tmpFn} which takes a single argument \texttt{xVec}.  the function should return the vector the values of the function $f(x)$ evaluated at the values in \texttt{xVec}.  
Hence plot the function $f(x)$ for $-3 < x <3$.
```{r}
tmpFn <- function(x)
{
  ifelse(x<0, x^2+2*x+3, ifelse(x<2, x+3, x^2+4*x-7))
}
tmp <- seq(-3, 3, len=100)
plot(tmp, tmpFn(tmp), type="l")
```

******

### 4. Write a function which takes a single argument which is a matrix.  The function should return a matrix which is the same as the function argument but every odd number is doubled.  
Hence the result of using the function on the matrix   

$$\begin{bmatrix}
1 & 1 & 3\\ 
5 & 2 & 6\\ 
-2 & -1 & -3
\end{bmatrix}$$

should be:

$$\begin{bmatrix}
2 & 2 & 6\\ 
10 & 2 & 6\\ 
-2 & -2 & -6
\end{bmatrix}$$
```{r}
tmpFn <- function(matrix)
{
  matrix[matrix%%2==1] <- 2*matrix[matrix%%2==1]
  matrix
}
```
******

### 5. Write a function which takes 2 arguements $n$ and $k$ which are positive integers.  It should return the $\mathit{n} x \mathit{n}$ matrix:


$$\begin{bmatrix}
k & 1 & 0 & 0 & \cdots  & 0 & 0\\ 
1 & k & 1 & 0 & \cdots  & 0 & 0\\
0 & 1 & k & 1 & \cdots  & 0 & 0\\
0 & 0 & 1 & k & \cdots  & 0 & 0\\
\cdot & \cdot & \cdot & \cdot & \cdot & \cdot & \cdot\\ 
0 & 0 & 0 & 0 & \cdots  & k & 1\\
0 & 0 & 0 & 0 & \cdots  & 1 & k
\end{bmatrix}$$
```{r}
tmpFn <- function(n, k)
{
  tmp <- diag(k, nrow=n)
  tmp[abs(row(tmp)-col(tmp))==1] <- 1
  tmp
}
```
******

### 6. Suppose an angle $\alpha$ is given as a positive real number of degrees.  
If $0 \leq \alpha < 90$ then it is quadrant 1.  If $90 \leq \alpha < 180$ then it is quadrant 2.  
if $180 \leq \alpha < 270$ then it is quadrant3.  if $270 \leq \alpha < 360$ then it is quadrant 4.  
if $360 \leq \alpha < 450$ then it is quadrant 1.  
And so on ...   

Write a function \texttt{quadrant(alpha)} which returns the quadrant of the angle $\alpha$.
```{r}
quadrant <- function(alpha)
{
  1+(alpha%%360)%%90
}
```
******

### 7.

(a) Zeller's congruence is the formula:

$$ f = ([2.6m - 0.2] + k + y + [y/4] + [c/4] - 2c)mod 7 $$
where $[x]$ denotes the integer part of $x$; for example $[7.5] = 7$.  

Zeller's congruence returns the day of the week $f$ given:

$k$ = the day of the month  
$y$ = the year in the century  
$c$ = the first 2 digits of the year (the century number)  
$m$ = the month number (where January is month 11 of the preceding year, February is month 12 of the preceding year, March is month 1, etc.)   
For example, the date 21/07/1`963 has $m = 5, k = 21, c = 19, y = 63$;  
the date 21/2/63 has $m=12, k=21, c=19, and y=62$.  
Write a function \texttt{weekday(day,month,year)} which returns the day of the week when given the numerical inputs of the day, month and year.  
Note that the value of 1 for $f$ denotes Sunday, 2 denotes Monday, etc.
```{r}
weekday <- function(day,month,year)
{
  month <- month-2
  if(month<=0){
    month <- month + 12
    year <- year -1
  }
  cc <- year%%100
  year <- year%%100
  tmp <- floor(2.6*month-.2)+day+year+year%%4+cc%%4-2*cc
}
```

(b) Does your function work if the input parameters \text{day, month, and year} are vectors with the same length and valid entries?
We need to remove the if statement for it to work with vectors
```{r}
weekday <- function(day, month, year)
{
  x <- month <=2
  month <- month-2+12*x
  year <- year-x
  cc <- year%%100
  year <- year%%100
  tmp <- floor(2.6*month-.2)+day+year+year%%4+cc%%4-2*cc
}
```
******

### 8. 

(a) Suppose $$x_{0}=1 \, and \, x_{1}=2$$ and $$x_{j}=x_{j-1}+\frac{2}{x_{j-1}} \, for\, j=1,2,.....$$ Write a function testLoop which takes the single argument n and returns the first n − 1 values of the
sequence $\left \{ x_{j} \right \}_{j\geq 0:}$ that means the values of $x_{0}, x_{1},x_{2},...,x_{n-2}.$
```{r}
testLoop <- function(n)
{
  xVec <- rep(NA, n-1)
  xVec[1] <- 1
  xVec[2] <- 2
  for(j in 3:(n-1))
    xVec[j] <- xVec[j-1]+2/xVec[j-1]
  xVec
  if( n<4 ) stop
}
```

(b) Now write a function testLoop2 which takes a single argument yVec which is a vector. The function
should return $\sum_{j=1}^{n}e^{j}$ where n is the length of yVec.
```{r}
testLoop2 <- function(yVec)
{
  n <- length(yVec)
  sum( exp(seq(along=yVec)) )
}
```
******

### 9. 
Solution of the difference equation $x_{n}=rx_{n-1}(1-x_{n-1})$, with starting value x1.
(a) Write a function quadmap( start, rho, niter ) which returns the vector $(x_{1},...,x_{n})$ where $x_{k}= rx_{k-1}(1-x_{k-1})$ and niter donates n, start donates x1, and rho donates r. 
```{r}
quadmap <- function(start, rho, niter)
{
  xVec <- rep(NA,niter)
  xVec[1] <- start
  for(i in 1:(niter-1)) {
    xVec[i + 1] <- rho * xVec[i] * (1 - xVec[i])
  }
  x 
}
```
(b) Now write a function which determines the number of iterations needed to get $\left | x_{n}-x_{n-1} \right |< 0.02$ So this function has only 2 arguments: start and rho. 
```{r}
quad2 <- function(start, rho, eps = 0.02)
{
  x1 <- start
  x2 <- rho*x1*(1-x1)
  niter <- 1
  while(abs(x1-x2) >= eps) {
    x1 <- x2
    x2 <- rho*x1*(1-x1)
    niter <- niter+1
  }
  niter
}
```
******

### 10. 

(a) Given a vector $(x_{1},...,x_{n})$ the sample autocorrelation of lag k is defined to be $r_{k}=\frac{\sum_{i=k+1}^{n}(x_{i}-\bar{x})(x_{i-k}-\bar{x})}{\sum_{i=1}^{n}(x_{i}-\bar{x})^{2}}$
Thus,
$r_{1}=\frac{\sum_{i=2}^{n}(x_{i}-\bar{x})(x_{i-1}-\bar{x})}{\sum_{i=1}^{n}(x_{i}-\bar{x})^{2}}=\frac{(x_{2}-\bar{x})(x_{1}-\bar{x})+...+(x_{n}-\bar{x})(x_{n-1}-\bar{x})}{\sum_{i=1}^{n}(x_{i}-\bar{x})^{2}}$ 
Write a function tmpFn(xVec) which takes a single argument xVec which is a vector and returns a list of two values: $r_{1}\, and\, r_{2}$. In particular, find $r_{1}\, and\, r_{2}$ for the vector (2,5,8,....53,56).
```{r}
tmpAcf <- function(xVec)
{
  xc <- xVec - mean(xVec)
  denom <- sum(xc^2)
  n <- length(x)
  r1 <- sum(xc[2:n]*xc[1:(n-1)])/denom
  r2 <- sum(xc[3:n]*xc[1:(n-2)])/denom
  list(r1=r1, r2=r2)
}
```

(b) (Harder.) Generalise the function so that it takes two arguments: the vector xVec and an integer k which lies between 1 and n − 1 where n is the length of xVec. The function should return the vector of the values $(r_{0}=1, r_{1},....,r_{k}).$
```{r}
tmpAcf <- function(x,k)
{
  xc <- x-mean(x)
  denom <- sum(xc^2)
  n <- length(x)
  tmpFn <- function(j){sum(xc[(j+1):n]*xc[1:(n-j)])/denom}
  c(1,sapply(1:k, tmpFn))
}
```

